Forecasting interrupted time series

### 2024-02-19 fits.qmd
Combined lit review with methods and discussion with conclusions

# Introduction

@@

It is important to note that in this study, we consider the problem of forecasting interrupted time series both during and after the event. This is a separate problem from change point detection [@truong2020selective] and anomaly detection [@talagala2020anomaly]. In the situations we investigate, we know that a change has occurred which has caused some unusual observations, and we want to forecast what will happen next. Change point detection refers to determining *when* a change occurred [@blazquez2021review], and anomaly detection aims to identify unusual observations with the intention of either minimising their value, or paying particular attention to them, and doing a root-cause investigation. Our study focus also differs from *intervention modeling*, sometimes referred to as *interrupted time series analysis and modelling* [@mcdowall2019interrupted], which is used when we have data about an outcome over time and the aim is to assess the impact of an intervention, policy, or program implemented at a specific time point. We are not trying to measure the impact of the interruption; rather we are trying to produce sensible forecasts during and after the interruption.

@@

We introduce several strategies to handling interruptions when forecasting in @sec-methods. We then apply these approaches to two real data sets in @sec-sec-examples. Finally, in @sec-conclusions, we discuss some of the advantages and disadvantages of each approach, and provide some recommendations for practitioners.

# Related works

The presence of unusual observations in time series analysis is a significant concern that has received attention in numerous studies. In this section, we provide a brief overview of some of these relevant publications.

One modelling approach that has gained popularity is the intervention model, also known as interrupted time series (ITS) modelling, introduced by @box1975intervention. This approach focuses on understanding how and whether outcomes change following the implementation of an intervention, policy, or programme at a specific time point. ITS models provide a robust tool for analysing policy and programme evaluation across diverse domains [@bernal2017interrupted;@mcdowall2019interrupted]. These models can forecast future values by developing models that incorporate a range of components, such as pre-intervention levels, trends, seasonality, covariates related to the intervention time, and other external factors that might affect the outcome variable. A related approach is the multiple regimes model, which is particularly useful in forecasting due to its ability to capture structural changes and transitions observed in interrupted time series data [@Koopg2007].

There are also statistical models that are used to detect changes in the time series data and that can model unusual observation and have proven effective in similar situations. A range of methods such as simple location shift and scale shift models [@aue2013structural], Bayesian [@barry1993bayesian], and supervised and supervised machine learning [@aminikhanghahi2017survey]. Facebook’s Prophet [@taylor2018forecasting] is an example of a time series forecasting method that can incorporate changepoint detection (automatically or manually), and adjust model parameters to capture the newest trends in a time series. Anomaly detection is another related area in time series analysis. In the context of anomaly detection, time series anomalies are defined as points or subsequences with unexpected or abnormal values, and the detection goal is to identify such anomalies [@blazquez2021review].

In the context of forecasting disruptions such as the COVID-19 pandemic, Bayesian approaches have been proposed for detecting changing points in outbreaks and forecasting case numbers under counterfactual scenarios [@dehning2020inferring]. Counterfactual analysis examines what would have happened to the forecast variable of interest if the disruptive event had not occurred. @athanasopoulos2023probabilistic proposed a variant of this approach for forecasting tourism recovery from COVID-19, combining forecast reconciliation and forecast combinations applied to historical data, to generate COVID-free counterfactual forecasts of what might have been if the pandemic never occurred. Then scenario-based judgemental probabilistic forecasts were compared with the counterfactual forecasts, to better understand the future recovery of the tourism sector from the pandemic.

Inventory stockouts are a common cause of data interruption in the supply chain. To deal with these interruptions, a popular technique is to provide an estimate for the observation corresponding to the stock-out period as if there were no stock-outs, and then generate projections appropriately. @Bell2000 presented an adjustment strategy for stock-out periods that involves smoothing demand volatility and correcting for stockouts using predicted variance conditioned on the observed stockout. @trapero2023demand also used the Tobit Kalman filter (TKF) for models presented in a state space framework for forecasting purposes. This method can efficiently deal with trends, seasonality, and exogenous influences in censored data, as long as the forecasting model functions within a state space framework.

Certain models offer flexibility in adjusting to changes in time series characteristics over time. Exponential Smoothing State Space models (ETS) represent an adaptive method that has demonstrated competitiveness in time series forecasting [@gardner2006exponential]. These approaches allow for the adaptation of model parameters over time to accommodate shifts in time series characteristics [@hyndman2002state]. Similarly, time-varying parameter models exhibit high adaptability [@harvey2006forecasting], enabling forecasting models to better accommodate interruptions in data over time. Time-varying parameters are frequently utilized in dynamic forecasting models, such as state-space models or Bayesian structural time series models [@talih2005structural]. Some models employ approaches to reduce the influence of affected observations during model training by assigning different weights. The utilization of weights extends beyond merely including or excluding observations and can also balance the degree of influence each observation has on the forecasting model [@khoshgoftaar2007empirical].

Disruptive events may potentially alter time series by causing gaps, or sometimes observations can be converted into missing data. While many time series forecasting models struggle with missing data, several approaches can internally handle missing data [@twala2008good]. @7157837 suggested a forecasting approach based on the Least Squares Support Vector Machine (LSSVM) that was particularly designed to handle time series forecasting with missing data. A different study by @tang2020joint focused on local and global temporal dynamics for multivariate time series forecasting with missing data. Their proposed framework uses a memory network to capture global temporal patterns with local data as key components.

Although several approaches exist, the literature lacks recommendations and comparisons of alternative strategies for forecasting time series data influenced by interruptions such as the COVID-19 pandemic. This paper aims to close this gap by proposing several methodologies for forecasting time series influenced by such disruptions.

@@

Highly adaptive models can adjust to the interruption as it is happening and will therefore be able to approximate the data generating process relatively well. For example, an ETS model with large smoothing parameters will be able to adjust to the interruption relatively quickly. This has the advantage of being a very simple solution that is easy to implement and fast to compute. There is no need to explicitly model the interruption, and so the model can be used for forecasting even if the timing or effect of the interruption is unknown.

However, the prediction intervals will be large because the model will have heavily discounted past data. In fact, the model will largely forget the past data other than the most recent observations, so there is no memory of the seasonal patterns and other dynamics that were present before the interruption. Consequently, the approach works best if there is no assumption that the post-interruption period will be similar to the pre-interruption period.

@@

A dynamic regression model with intervention covariates can be used to model the interruption explicitly. For example, if the intervention involves a simple level shift with a reverse level shift at the end of the intervention, we can use a dummy variable to indicate the interruption period and allow the model to adjust to the interruption. More complicated interventions can be handled by using more covariates.

@@

Many time series models will handle periods of missing values. So the problematic observations that occur during the period of disruption can be set to missing, and the model should continue to produce forecasts as if the interruption had not occurred. Of course, the forecasts will not be accurate for the period of disruption, but they can be interpreted as "what might have been". This solution requires a judgement to be made about when the disruption has begun, and when normality resumes.

Because no information is retained during the disruption, the prediction intervals will become large during the disruption, and after the disruption, they will remain large until the model has enough data to estimate the forecast distribution more accurately.

@@

@@

# Discussion

@@

# Conclusion

>>>>

# Introduction {#sec-introduction}

@@

Although several approaches to the problem exist, the literature lacks recommendations and comparisons of alternative strategies for forecasting time series data influenced by interruptions such as the COVID-19 pandemic. This paper aims to close this gap by proposing several methodologies for forecasting time series influenced by such disruptions.

It is important to note that in this study, we consider the problem of forecasting interrupted time series both during and after the event. This is a separate problem from change point detection [@aue2013structural;@truong2020selective] and anomaly detection [@talagala2020anomaly]. In the situations we investigate, we know that a change has occurred which has caused some unusual observations, and we want to forecast what will happen next. Change point detection refers to determining *when* a change occurred [@blazquez2021review], and anomaly detection aims to identify unusual observations with the intention of either minimising their value, or paying particular attention to them, and doing a root-cause investigation. Our study focus also differs from *intervention modeling*, sometimes referred to as *interrupted time series analysis and modelling* [@mcdowall2019interrupted], which is used when we have data about an outcome over time and the aim is to assess the impact of an intervention, policy, or program implemented at a specific time point. We are not trying to measure the impact of the interruption; rather we are trying to produce sensible forecasts during and after the interruption.

@@

We introduce several strategies to handling interruptions when forecasting in @sec-methods. We then apply these approaches to two real data sets in @sec-examples. Finally, in @sec-conclusions, we discuss some of the advantages and disadvantages of each approach, and provide some recommendations for practitioners.

@@

Highly adaptive models offer flexibility in adjusting to changes in time series characteristics over time. These models can adjust to the interruption as it is happening and will therefore be able to approximate the data generating process relatively well.

Exponential Smoothing State Space models (ETS) represent an adaptive method that has demonstrated competitiveness in time series forecasting [@gardner2006exponential]. These models allow for the adaptation of model parameters over time to accommodate shifts in time series characteristics [@hyndman2002state]. For example, an ETS model with large smoothing parameters will be able to adjust to the interruption relatively quickly. This has the advantage of being a very simple solution that is easy to implement and fast to compute. There is no need to explicitly model the interruption, and so the model can be used for forecasting even if the timing or effect of the interruption is unknown.

Similarly, time-varying parameter models exhibit high adaptability [@harvey2006forecasting], enabling forecasting models to better accommodate interruptions in data over time. Time-varying parameters are frequently utilized in dynamic forecasting models, such as state-space models or Bayesian structural time series models [@talih2005structural]. Some models employ approaches to reduce the influence of affected observations during model training by assigning different weights. The utilization of weights extends beyond merely including or excluding observations and can also balance the degree of influence each observation has on the forecasting model [@khoshgoftaar2007empirical].

The prediction intervals from highly adaptive models tend to be wide because the model will have heavily discounted past data. In fact, the model will largely forget the past data other than the most recent observations, so there is no memory of the seasonal patterns and other dynamics that were present before the interruption. Consequently, the approach works best if there is no assumption that the post-interruption period will be similar to the pre-interruption period.

@@

One modelling approach that has gained popularity in handling disruptions is the intervention model, also known as interrupted time series (ITS) modelling, introduced by @box1975intervention. This approach focuses on understanding how and whether outcomes change following the implementation of an intervention, policy, or programme at a specific time point. ITS models provide a robust tool for analysing policy and programme evaluation across diverse domains [@bernal2017interrupted;@mcdowall2019interrupted]. These models can forecast future values by developing models that incorporate a range of components, such as pre-intervention levels, trends, seasonality, covariates related to the intervention time, and other external factors that might affect the outcome variable. A related approach is the multiple regimes model, which is particularly useful in forecasting due to its ability to capture structural changes and transitions observed in interrupted time series data [@Koopg2007].

A special case of the intervention models proposed by @box1975intervention is a dynamic regression (or regression with ARIMA errors) model. This model includes covariates describing the interruption explicitly, with an ARIMA process describing the remaaining time series dynamics. For example, if the intervention involves a simple level shift with a reverse level shift at the end of the intervention, we can use a dummy variable to indicate the interruption period and allow the model to adjust to the interruption. More complicated interventions can be handled by using more covariates.

@@

Disruptive events may potentially alter time series by causing gaps. Alternatively, problematic observations that occur during the period of disruption can be set to missing, and a model fitted to the remaining observations should continue to produce forecasts as if the interruption had not occurred. Of course, the forecasts will not be accurate for the period of disruption, but they can be interpreted as "what might have been". This solution requires a judgement to be made about when the disruption has begun, and when normality resumes.

While many time series forecasting models struggle with missing data, several approaches can internally handle missing data [@twala2008good]. @7157837 suggested a forecasting approach based on a Least Squares Support Vector Machine (LSSVM) that was particularly designed to handle time series forecasting with missing data. A different study by @tang2020joint focused on local and global temporal dynamics for multivariate time series forecasting with missing data. Their proposed framework uses a memory network to capture global temporal patterns with local data as key components.

In this approach, no information is retained during the disruption, so the prediction intervals will become large during the disruption, and after the disruption, they will remain large until the model has enough data to estimate the forecast distribution more accurately.

@@

These approaches are sometimes called "counterfactual" methods, as they examine what would have happened to the forecast variable of interest if the disruptive event had not occurred. @athanasopoulos2023probabilistic proposed a variant of this approach for forecasting tourism recovery from COVID-19, combining forecast reconciliation and forecast combinations applied to historical data, to generate COVID-free counterfactual forecasts of what might have been if the pandemic never occurred. Then scenario-based judgemental probabilistic forecasts were compared with the counterfactual forecasts, to better understand the future recovery of the tourism sector from the pandemic.

Inventory stockouts are a common cause of data interruption in the supply chain. To deal with these interruptions, a counterfactual approach is popular, whereby the demand corresponding to the stock-out period is estimated as if there were no stock-outs, and then forecasts are obtained. @Bell2000 presented an adjustment strategy for stock-out periods that involves smoothing demand volatility and correcting for stockouts using predicted variance conditioned on the observed stockout. @trapero2023demand also used the Tobit Kalman filter (TKF) for models presented in a state space framework for forecasting purposes. This method can efficiently deal with trends, seasonality, and exogenous influences in censored data, as long as the forecasting model functions within a state space framework.

In the context of forecasting disruptions such as the COVID-19 pandemic, Bayesian approaches have also been proposed for detecting changing points in outbreaks and forecasting case numbers under counterfactual scenarios [@dehning2020inferring].

@@

# Discussion and conclusions {#sec-conclusions}

@@

### 2024-02-19 fits.tex
Combined lit review with methods and discussion with conclusions

\setstretch{1.5}
\section{Introduction}\label{introduction}

@@

It is important to note that in this study, we consider the problem of
forecasting interrupted time series both during and after the event.
This is a separate problem from change point detection
\autocite{truong2020selective} and anomaly detection
\autocite{talagala2020anomaly}. In the situations we investigate, we
know that a change has occurred which has caused some unusual
observations, and we want to forecast what will happen next. Change
point detection refers to determining \emph{when} a change occurred
\autocite{blazquez2021review}, and anomaly detection aims to identify
unusual observations with the intention of either minimising their
value, or paying particular attention to them, and doing a root-cause
investigation. Our study focus also differs from \emph{intervention
modeling}, sometimes referred to as \emph{interrupted time series
analysis and modelling} \autocite{mcdowall2019interrupted}, which is
used when we have data about an outcome over time and the aim is to
assess the impact of an intervention, policy, or program implemented at
a specific time point. We are not trying to measure the impact of the
interruption; rather we are trying to produce sensible forecasts during
and after the interruption.

@@

We introduce several strategies to handling interruptions when
forecasting in Section~\ref{sec-methods}. We then apply these approaches
to two real data sets in \textbf{?@sec-sec-examples}. Finally, in
\textbf{?@sec-conclusions}, we discuss some of the advantages and
disadvantages of each approach, and provide some recommendations for
practitioners.

\section{Related works}\label{related-works}

The presence of unusual observations in time series analysis is a
significant concern that has received attention in numerous studies. In
this section, we provide a brief overview of some of these relevant
publications.

One modelling approach that has gained popularity is the intervention
model, also known as interrupted time series (ITS) modelling, introduced
by \textcite{box1975intervention}. This approach focuses on
understanding how and whether outcomes change following the
implementation of an intervention, policy, or programme at a specific
time point. ITS models provide a robust tool for analysing policy and
programme evaluation across diverse domains
\autocite{bernal2017interrupted,mcdowall2019interrupted}. These models
can forecast future values by developing models that incorporate a range
of components, such as pre-intervention levels, trends, seasonality,
covariates related to the intervention time, and other external factors
that might affect the outcome variable. A related approach is the
multiple regimes model, which is particularly useful in forecasting due
to its ability to capture structural changes and transitions observed in
interrupted time series data \autocite{Koopg2007}.

There are also statistical models that are used to detect changes in the
time series data and that can model unusual observation and have proven
effective in similar situations. A range of methods such as simple
location shift and scale shift models \autocite{aue2013structural},
Bayesian \autocite{barry1993bayesian}, and supervised and supervised
machine learning \autocite{aminikhanghahi2017survey}. Facebook's Prophet
\autocite{taylor2018forecasting} is an example of a time series
forecasting method that can incorporate changepoint detection
(automatically or manually), and adjust model parameters to capture the
newest trends in a time series. Anomaly detection is another related
area in time series analysis. In the context of anomaly detection, time
series anomalies are defined as points or subsequences with unexpected
or abnormal values, and the detection goal is to identify such anomalies
\autocite{blazquez2021review}.

In the context of forecasting disruptions such as the COVID-19 pandemic,
Bayesian approaches have been proposed for detecting changing points in
outbreaks and forecasting case numbers under counterfactual scenarios
\autocite{dehning2020inferring}. Counterfactual analysis examines what
would have happened to the forecast variable of interest if the
disruptive event had not occurred.
\textcite{athanasopoulos2023probabilistic} proposed a variant of this
approach for forecasting tourism recovery from COVID-19, combining
forecast reconciliation and forecast combinations applied to historical
data, to generate COVID-free counterfactual forecasts of what might have
been if the pandemic never occurred. Then scenario-based judgemental
probabilistic forecasts were compared with the counterfactual forecasts,
to better understand the future recovery of the tourism sector from the
pandemic.

Inventory stockouts are a common cause of data interruption in the
supply chain. To deal with these interruptions, a popular technique is
to provide an estimate for the observation corresponding to the
stock-out period as if there were no stock-outs, and then generate
projections appropriately. \textcite{Bell2000} presented an adjustment
strategy for stock-out periods that involves smoothing demand volatility
and correcting for stockouts using predicted variance conditioned on the
observed stockout. \textcite{trapero2023demand} also used the Tobit
Kalman filter (TKF) for models presented in a state space framework for
forecasting purposes. This method can efficiently deal with trends,
seasonality, and exogenous influences in censored data, as long as the
forecasting model functions within a state space framework.

Certain models offer flexibility in adjusting to changes in time series
characteristics over time. Exponential Smoothing State Space models
(ETS) represent an adaptive method that has demonstrated competitiveness
in time series forecasting \autocite{gardner2006exponential}. These
approaches allow for the adaptation of model parameters over time to
accommodate shifts in time series characteristics
\autocite{hyndman2002state}. Similarly, time-varying parameter models
exhibit high adaptability \autocite{harvey2006forecasting}, enabling
forecasting models to better accommodate interruptions in data over
time. Time-varying parameters are frequently utilized in dynamic
forecasting models, such as state-space models or Bayesian structural
time series models \autocite{talih2005structural}. Some models employ
approaches to reduce the influence of affected observations during model
training by assigning different weights. The utilization of weights
extends beyond merely including or excluding observations and can also
balance the degree of influence each observation has on the forecasting
model \autocite{khoshgoftaar2007empirical}.

Disruptive events may potentially alter time series by causing gaps, or
sometimes observations can be converted into missing data. While many
time series forecasting models struggle with missing data, several
approaches can internally handle missing data \autocite{twala2008good}.
\textcite{7157837} suggested a forecasting approach based on the Least
Squares Support Vector Machine (LSSVM) that was particularly designed to
handle time series forecasting with missing data. A different study by
\textcite{tang2020joint} focused on local and global temporal dynamics
for multivariate time series forecasting with missing data. Their
proposed framework uses a memory network to capture global temporal
patterns with local data as key components.

Although several approaches exist, the literature lacks recommendations
and comparisons of alternative strategies for forecasting time series
data influenced by interruptions such as the COVID-19 pandemic. This
paper aims to close this gap by proposing several methodologies for
forecasting time series influenced by such disruptions.

@@

Highly adaptive models can adjust to the interruption as it is happening
and will therefore be able to approximate the data generating process
relatively well. For example, an ETS model with large smoothing
parameters will be able to adjust to the interruption relatively
quickly. This has the advantage of being a very simple solution that is
easy to implement and fast to compute. There is no need to explicitly
model the interruption, and so the model can be used for forecasting
even if the timing or effect of the interruption is unknown.

However, the prediction intervals will be large because the model will
have heavily discounted past data. In fact, the model will largely
forget the past data other than the most recent observations, so there
is no memory of the seasonal patterns and other dynamics that were
present before the interruption. Consequently, the approach works best
if there is no assumption that the post-interruption period will be
similar to the pre-interruption period.

@@

A dynamic regression model with intervention covariates can be used to
model the interruption explicitly. For example, if the intervention
involves a simple level shift with a reverse level shift at the end of
the intervention, we can use a dummy variable to indicate the
interruption period and allow the model to adjust to the interruption.
More complicated interventions can be handled by using more covariates.

@@

Many time series models will handle periods of missing values. So the
problematic observations that occur during the period of disruption can
be set to missing, and the model should continue to produce forecasts as
if the interruption had not occurred. Of course, the forecasts will not
be accurate for the period of disruption, but they can be interpreted as
``what might have been''. This solution requires a judgement to be made
about when the disruption has begun, and when normality resumes.

Because no information is retained during the disruption, the prediction
intervals will become large during the disruption, and after the
disruption, they will remain large until the model has enough data to
estimate the forecast distribution more accurately.

@@

@@

\section{Discussion}\label{discussion}

@@

\section{Conclusion}\label{conclusion}

>>>>

\setstretch{1.5}
\section{Introduction}\label{sec-introduction}

@@

Although several approaches to the problem exist, the literature lacks
recommendations and comparisons of alternative strategies for
forecasting time series data influenced by interruptions such as the
COVID-19 pandemic. This paper aims to close this gap by proposing
several methodologies for forecasting time series influenced by such
disruptions.

It is important to note that in this study, we consider the problem of
forecasting interrupted time series both during and after the event.
This is a separate problem from change point detection
\autocite{aue2013structural,truong2020selective} and anomaly detection
\autocite{talagala2020anomaly}. In the situations we investigate, we
know that a change has occurred which has caused some unusual
observations, and we want to forecast what will happen next. Change
point detection refers to determining \emph{when} a change occurred
\autocite{blazquez2021review}, and anomaly detection aims to identify
unusual observations with the intention of either minimising their
value, or paying particular attention to them, and doing a root-cause
investigation. Our study focus also differs from \emph{intervention
modeling}, sometimes referred to as \emph{interrupted time series
analysis and modelling} \autocite{mcdowall2019interrupted}, which is
used when we have data about an outcome over time and the aim is to
assess the impact of an intervention, policy, or program implemented at
a specific time point. We are not trying to measure the impact of the
interruption; rather we are trying to produce sensible forecasts during
and after the interruption.

@@

We introduce several strategies to handling interruptions when
forecasting in Section~\ref{sec-methods}. We then apply these approaches
to two real data sets in Section~\ref{sec-examples}. Finally, in
Section~\ref{sec-conclusions}, we discuss some of the advantages and
disadvantages of each approach, and provide some recommendations for
practitioners.

@@

Highly adaptive models offer flexibility in adjusting to changes in time
series characteristics over time. These models can adjust to the
interruption as it is happening and will therefore be able to
approximate the data generating process relatively well.

Exponential Smoothing State Space models (ETS) represent an adaptive
method that has demonstrated competitiveness in time series forecasting
\autocite{gardner2006exponential}. These models allow for the adaptation
of model parameters over time to accommodate shifts in time series
characteristics \autocite{hyndman2002state}. For example, an ETS model
with large smoothing parameters will be able to adjust to the
interruption relatively quickly. This has the advantage of being a very
simple solution that is easy to implement and fast to compute. There is
no need to explicitly model the interruption, and so the model can be
used for forecasting even if the timing or effect of the interruption is
unknown.

Similarly, time-varying parameter models exhibit high adaptability
\autocite{harvey2006forecasting}, enabling forecasting models to better
accommodate interruptions in data over time. Time-varying parameters are
frequently utilized in dynamic forecasting models, such as state-space
models or Bayesian structural time series models
\autocite{talih2005structural}. Some models employ approaches to reduce
the influence of affected observations during model training by
assigning different weights. The utilization of weights extends beyond
merely including or excluding observations and can also balance the
degree of influence each observation has on the forecasting model
\autocite{khoshgoftaar2007empirical}.

The prediction intervals from highly adaptive models tend to be wide
because the model will have heavily discounted past data. In fact, the
model will largely forget the past data other than the most recent
observations, so there is no memory of the seasonal patterns and other
dynamics that were present before the interruption. Consequently, the
approach works best if there is no assumption that the post-interruption
period will be similar to the pre-interruption period.

@@

One modelling approach that has gained popularity in handling
disruptions is the intervention model, also known as interrupted time
series (ITS) modelling, introduced by \textcite{box1975intervention}.
This approach focuses on understanding how and whether outcomes change
following the implementation of an intervention, policy, or programme at
a specific time point. ITS models provide a robust tool for analysing
policy and programme evaluation across diverse domains
\autocite{bernal2017interrupted,mcdowall2019interrupted}. These models
can forecast future values by developing models that incorporate a range
of components, such as pre-intervention levels, trends, seasonality,
covariates related to the intervention time, and other external factors
that might affect the outcome variable. A related approach is the
multiple regimes model, which is particularly useful in forecasting due
to its ability to capture structural changes and transitions observed in
interrupted time series data \autocite{Koopg2007}.

A special case of the intervention models proposed by
\textcite{box1975intervention} is a dynamic regression (or regression
with ARIMA errors) model. This model includes covariates describing the
interruption explicitly, with an ARIMA process describing the remaaining
time series dynamics. For example, if the intervention involves a simple
level shift with a reverse level shift at the end of the intervention,
we can use a dummy variable to indicate the interruption period and
allow the model to adjust to the interruption. More complicated
interventions can be handled by using more covariates.

@@

Disruptive events may potentially alter time series by causing gaps.
Alternatively, problematic observations that occur during the period of
disruption can be set to missing, and a model fitted to the remaining
observations should continue to produce forecasts as if the interruption
had not occurred. Of course, the forecasts will not be accurate for the
period of disruption, but they can be interpreted as ``what might have
been''. This solution requires a judgement to be made about when the
disruption has begun, and when normality resumes.

While many time series forecasting models struggle with missing data,
several approaches can internally handle missing data
\autocite{twala2008good}. \textcite{7157837} suggested a forecasting
approach based on a Least Squares Support Vector Machine (LSSVM) that
was particularly designed to handle time series forecasting with missing
data. A different study by \textcite{tang2020joint} focused on local and
global temporal dynamics for multivariate time series forecasting with
missing data. Their proposed framework uses a memory network to capture
global temporal patterns with local data as key components.

In this approach, no information is retained during the disruption, so
the prediction intervals will become large during the disruption, and
after the disruption, they will remain large until the model has enough
data to estimate the forecast distribution more accurately.

@@

These approaches are sometimes called ``counterfactual'' methods, as
they examine what would have happened to the forecast variable of
interest if the disruptive event had not occurred.
\textcite{athanasopoulos2023probabilistic} proposed a variant of this
approach for forecasting tourism recovery from COVID-19, combining
forecast reconciliation and forecast combinations applied to historical
data, to generate COVID-free counterfactual forecasts of what might have
been if the pandemic never occurred. Then scenario-based judgemental
probabilistic forecasts were compared with the counterfactual forecasts,
to better understand the future recovery of the tourism sector from the
pandemic.

Inventory stockouts are a common cause of data interruption in the
supply chain. To deal with these interruptions, a counterfactual
approach is popular, whereby the demand corresponding to the stock-out
period is estimated as if there were no stock-outs, and then forecasts
are obtained. \textcite{Bell2000} presented an adjustment strategy for
stock-out periods that involves smoothing demand volatility and
correcting for stockouts using predicted variance conditioned on the
observed stockout. \textcite{trapero2023demand} also used the Tobit
Kalman filter (TKF) for models presented in a state space framework for
forecasting purposes. This method can efficiently deal with trends,
seasonality, and exogenous influences in censored data, as long as the
forecasting model functions within a state space framework.

In the context of forecasting disruptions such as the COVID-19 pandemic,
Bayesian approaches have also been proposed for detecting changing
points in outbreaks and forecasting case numbers under counterfactual
scenarios \autocite{dehning2020inferring}.

@@

\section{Discussion and conclusions}\label{sec-conclusions}

@@
